---
title: "Leaflet Demonstration Project"
date: "February 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

#### Project Summary

This project is a simple demonstration of the interactive map package for R, Leaflet. It uses National Park visitation data for select national parks in the continental United States. The project embeds an interactive Leaflet plot in a Shiny App to provide further interactivity in the form of a slider bar that allows users to select a year.


```{r  echo=FALSE, }
# load packages
library(ggmap)
library(leaflet)
library(readr)
library(reshape2)
library(rgdal)
library(shiny)
```


```{r get base data, echo=FALSE}

# Load Parks Visitorship data
parks1 <- read_csv("annual_park_visitation_1979_1996.csv")
parks2 <- read_csv("annual_park_visitation_1997_2016.csv")

# Melt data
parks1_melted <- melt(parks1, id.vars=c("park_name"), 
                      variable.name="year",
                      value.name="visits")

parks2_melted <- melt(parks2, id.vars=c("park_name"), 
                      variable.name="year",
                      value.name="visits")

parks <- rbind(parks1_melted, parks2_melted)  
```

```{r format clean data, echo=FALSE}

# Remove repeat and missing-values observations
parks <- subset(parks, !is.na(visits))

# Subset parks on National Parks and National Monuments only
parks <- subset(parks, grepl("NP\\b", park_name) & 
                  !grepl("&", park_name))

# Expand park designations
parks$park_name <- sapply(parks$park_name, function(name){
  gsub("NP", "National Park", name)
})

# Year to class integer
parks$year <- as.integer(as.character(parks$year))
```

```{r get geo-data, echo=FALSE}

# Load dataframe with park coordinates
park_coords_df <- read_csv("parks_coords.csv")

# Subset on national parks and monuments
park_coords_df <- subset(park_coords_df, 
                         grepl("National Park\\b", park_name) &
                         !grepl("&", park_name))

# Subset for continental US
park_coords_df <- subset(park_coords_df, 
                         lon > -125 & lon < 65 &
                           lat > 25 & lat < 50)

```

```{r merge data, echo=FALSE}

# Merge datasets, some parks in park visitation dataset dropped bc not in coord df
parks_full <- merge(parks, park_coords_df, by="park_name")

# Park code to uppercase
parks_full$park_number <- toupper(parks_full$park_number)

```

```{r get park boundary shapes, echo=FALSE}

# Load shape files for park boundaries
park_bounds <- readOGR("nps_boundary.shp", layer = "nps_boundary",
                       verbose = FALSE)

# Subset boundary files on selected parks and monuments
park_bounds <- subset(park_bounds, UNIT_CODE %in% parks_full$park_number)
```


## Annual National Park Visitation by Year

```{r user interface, echo=FALSE}

# Define UI for application 
ui <- fluidPage(leafletOutput("parkMap"),
                      absolutePanel(bottom = 0,
                                    left = 25,
                                    draggable = TRUE,
                                    sliderInput("years",
                                                "Year of Data Summary:",
                                                min = min(parks_full$year),
                                                max = max(parks_full$year),
                                                value = min(parks_full$year),
                                                sep = "",
                                                step = 1,
                                                animate=TRUE)))


```

```{r server, echo=FALSE}

# Define required server logic 
server <- function(input, output){

  # Subset data on selected year
  parks_react <- reactive({
    parks_full[parks_full$year==input$years,]
  }) 
  
  #Build base map
  output$parkMap <- renderLeaflet({
    
  # Instantiate and build leaflet app
        parks_full%>% leaflet()%>%
          addTiles()%>%
          addPolygons(color = "grey50", weight = 1, smoothFactor = 0.5,
                  opacity = 1.0, fillOpacity = 0.5,
                  fillColor = "blue", data = park_bounds)%>%
          fitBounds(lng1=-124.848974, lat1=24.396308, 
                   lng2=-66.885444, lat2=49.384358)

  })
    
  observe({
    # Add annual plot data to base map
    
    # Get base map
    leafletProxy("parkMap", data=parks_react())%>%
      
    # Clear any previously plotted markers
    clearMarkers()%>%
      
    # Plot new markers by selected year
    addCircleMarkers(radius = ~visits/150000+10, 
                     color = "purple",
                     popup = paste("<center>", 
                                   parks_react()$park_name, "<br>",
                                   "Visitors in ", parks_react()$year, ": ", 
                                   format(parks_react()$visits, big.mark = ","), 
                                   "<br>",
                                   "</center"))
  })
}

# Run the application 
shinyApp(ui = ui, server = server)

